Multiscalar models for systems of cities

### 2023-09-14 Docs/Communications/2023/ECTQG2023/slides.tex
slides

\date{ECTQG 2023\\\smallskip
September 15th 2023
}

@@

\begin{center}
	\begin{columns}
	\begin{column}{0.45\linewidth}
	\includegraphics[width=\textwidth]{../../2020/ALife2020/figures/evoltheory_scales.png}
	\end{column}
	\begin{column}{0.22\linewidth}
	\includegraphics[width=\textwidth]{../../2020/ALife2020/figures/simpop1.png}\\\medskip
	\includegraphics[width=\textwidth]{../../2020/ALife2020/figures/setup_synth_1_tick100.png}
	\end{column}
	\end{columns}
\end{center}

\footnotesize
\textit{An evolutionary urban theory considering cities as systems within systems of cities \cite{pumain2018evolutionary}; Simpop 1 model \cite{sanders1997simpop}; SimpopNet model \cite{schmitt2014modelisation}}

@@

\sframe{Multi-modeling urban systems dynamics}{

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

\date{ECTQG 2023\\\smallskip
Session Theoretical Geography 1\\\smallskip
September 15th 2023
}

@@

\begin{center}
	\begin{columns}
	\begin{column}{0.49\linewidth}
	\includegraphics[width=\textwidth]{../../2020/ALife2020/figures/evoltheory_scales.png}
	\end{column}
	\begin{column}{0.3\linewidth}
	\includegraphics[width=\textwidth]{../../2020/ALife2020/figures/simpop1.png}\\\medskip
	\includegraphics[width=\textwidth]{../../2020/ALife2020/figures/setup_synth_1_tick100.png}
	\end{column}
	\end{columns}
\end{center}

%\footnotesize
An evolutionary urban theory considering cities as systems within systems of cities \cite{pumain2018evolutionary}; Simpop 1 model \cite{sanders1997simpop}; SimpopNet model \cite{schmitt2014modelisation}

@@

\sframe{Multi-modeling urban systems dynamics for sustainability}{

@@

\begin{center}
	\includegraphics[width=0.51\textwidth]{../../../Papers/SDGTradeoffs/SDGTradeoffs-paper/figures/Fig3.png}
	\includegraphics[width=0.42\textwidth]{../../../../../SDGTradeoffs/Docs/Communications/2022/FRCCS2022/figures/radar_sdgs.png}
\end{center}

Trade-offs between sustainable development goals in systems of cities \cite{raimbault2020empowering}; multi-modeling for many-objectives SDGs optimisation \cite{raimbault2022trade}

@@

\begin{center}
	\includegraphics[width=0.49\textwidth]{../../../../../CityNetwork/Docs/ThesisMemoire/Final/Figures/Final/7-3-3-fig-lutecia-governance.jpg}
	\includegraphics[width=0.49\textwidth]{../../../../../NetworksTerritories/CoevolutionNwTerritories/Docs/Papers/HierarchyCoevolution/chapter/figuresraw/ex_realphysical_1975_tf.png}
\end{center}

Simulation models for systems of cities and transportation networks making a slight bridge between scales: transportation governance modeling \cite{raimbault2021introducing}; physical network self-reinforcment \cite{raimbault2020hierarchy}

@@

$\rightarrow$ multi-scale simulation models are needed to explore sustainable policies \cite{rozenblat2018conclusion}

$\rightarrow$ multi-scale urban models remain however underexplored in the literature

\bigskip
\bigskip

\textbf{This contribution: } \textit{to synthesise lessons from two different recent tentatives to build multi-scale urban models with a strong coupling between scales: an urban morphogenesis model \cite{raimbault2021strong} and an innovation diffusion model \cite{raimbault2023innovation}, learnt from the systematic exploration of these models using OpenMOLE \cite{reuillon2013openmole}}

@@

\begin{center}
\includegraphics[width=\textwidth]{../../../../../UrbanGrowth/Docs/Papers/Multiscale/paper/figures/Fig1.jpg}
\end{center}

A strong coupling explored by \cite{raimbault2021strong} between urban morphogenesis agregation-diffusion processes at the meso scale \cite{raimbault2018calibration} and interactions between cities at the macro scale \cite{raimbault2020indirect}

@@

\begin{center}
\includegraphics[width=0.58\textwidth]{../../../../../UrbanGrowth/Docs/Papers/Multiscale/paper/figures/Fig5.jpg}
\includegraphics[width=0.4\textwidth]{../../../../../UrbanGrowth/Docs/Papers/Multiscale/paper/figures/Fig7.jpg}
\end{center}

(Left) Non-monotonous effect of policy parameters linking scales; (Right) Bi-objective optimisation of contradictory objectives across scales

@@

\begin{center}
	\includegraphics[width=0.75\textwidth]{../../../Papers/InnovationMultiscale/InnovationMultiscale-paper/slides/figures/model_4.png}
\end{center}

A bi-scale model explored by \cite{raimbault2023innovation}, coupling urban evolution based on innovation diffusion at the macro scale \cite{raimbault2020model}, with intra-urban innovation cluster dynamics between firms for the emergence of new innovations \cite{raimbault2022innovation}; quantification of strong emergence using indicators by \cite{rosas2020reconciling}

@@

\begin{center}
\includegraphics[width=0.49\textwidth]{../../../Papers/InnovationMultiscale/InnovationMultiscale-paper/figures/paretoDiversity-Fitness_colordG_sizetheta.png}
\includegraphics[width=0.49\textwidth]{../../../Papers/InnovationMultiscale/InnovationMultiscale-paper/figures/pse-psi-delta-utility_colorGamma_sizetheta.png}
\end{center}

(Left) bi-optimisation between contradictory objectives across scales; (Right) diversity of emergence regimes obtained with a diversity search algorithm

@@

$\rightarrow$ strong emergence quantified in a high diversity of regimes for the innovation diffusion model, probable (given cross-scale parameter influence, but not quantified) for the urban morphogenesis model

$\rightarrow$ optimisation objectives across scales seem generally contradictory, and can be simultaneously optimised with Pareto trade-offs
